---
title: "Exercises 01"
author: "Bjarne Taulo Sørensen"
date: "Week 43 - 2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




<br />\    
  

<!-- Your own secret notes can be written here between the HTML signs -->   



```{r bibliotek, eval=TRUE, include=TRUE}


if (!dir.exists("data")) dir.create("data")

# Gem dine datasæt i dette bibliotek

```


```{r pacman, eval=TRUE, include=FALSE}

pacman::p_load("tidyverse", "magrittr", "nycflights13", "gapminder",
               "Lahman", "maps", "lubridate", "pryr", "hms", "hexbin",
               "feather", "htmlwidgets", "broom", "pander", "modelr",
               "XML", "httr", "jsonlite", "lubridate", "microbenchmark",
               "splines", "ISLR", "MASS", "testthat", "leaps", "caret",
               "RSQLite", "class", "babynames", "nasaweather",
               "fueleconomy", "viridis", "boot", "devtools")

```

<br />\   

<!-- Your own secret notes can be written here between the HTML signs -->

# Problem 1

Make a new data set (call it Extra001) based on “data20171001”. Delete observations where “Department” has a missing value or the value “-“. Delete columns that are constants. 

Remember to make comments.

```{r problem 1 solution, eval=FALSE, include=FALSE}


Canteen_data_inRformat <- readRDS(file="data/data20171001.Rda")
Canteen_data_inRformat

View(Canteen_data_inRformat)


table(Canteen_data_inRformat$Units01)

Extra001 <- Canteen_data_inRformat %>%
 filter(!Department %in% c(NA, " - ", "-", " ", "")) %>%
  dplyr::select(-Segment, -Units01) %>% 
  mutate(Department = trimws(Department))

# Better:
Extra001 <- Canteen_data_inRformat %>%
 filter(!trimws(Department) %in% c(NA, "-", "")) %>%
  dplyr::select(-Segment, -Units01) 

View(Extra001)

```

# Problem 2

Make an id-variable that has the value 1 in line one and 2 in line 2 etc. (Save Extra001 in your working directory in R format).

Remember to make comments.


```{r problem 2, eval=FALSE, include=FALSE}
Extra001 <- Canteen_data_inRformat %>%
 filter(!trimws(Department) %in% c(NA, "-", "")) %>%
  dplyr::select(-Segment, -Units01) %>%
   mutate(id = row_number())

saveRDS(Extra001, file = "data/Extra001.Rda")


View(Extra001)

```

# Problem 3


Make a new data set (call it Extra002) based on Extra001. Delete observations where “Subsidy” has a value, and where “Price01” is negative.

Remember to make comments.
  
```{r problem 3, eval=FALSE, include=FALSE}

Extra002 <- Extra001 %>% 
  filter(!(!Subsidy %in% c(" ") & Price01 < 0))  #Filter the subsidies 
 # %>%  filter(Price01 > 0) #Filter the rest i.e. the bottle deposits, if that actually was the plan
View(Extra002)
 
```


# Problem 4


Make a new data set (call it Extra003) based on Extra001. Keep only observations where “Subsidy” has a value. Check that all values of “Price01” are negative. Make an extra id variable that is equal to the first one you made minus one. 

Remember to make comments.


```{r problem 4, eval=FALSE, include=FALSE}

Extra003 <- Extra001 %>% 
 filter(!Subsidy %in% c(" ")) %>% 
  mutate(id_extra = as.integer(id - 1)) %>%
   dplyr::select(-Date01, -Timofday03, -(Department:Product), -id) %>%  # Problem 5
    rename(Initials_extra = Initials, Sub_extra = Subsidy,  Price_extra = Price01)  # Problem 5

View(Extra003)

Extra003 %>% 
  summarise(max(Price_extra))

max(Extra003$Price_extra)


```

# Problem 5


Delete the following variables: “Date01”, “Timofday03”, “Department”:”Product”, “Units01”, and the first id-variable you made. Rename the rest of the variables except the new id-variable. 

Remember to make comments.


```{r problem 9, eval=FALSE, include=FALSE}

# See Problem 4

```

# Problem 6

Merge Extra002 and Extra003, and sum Price_extra and Price01.

Remember to make comments.  


```{r problem 6 solution, eval=FALSE, include=FALSE}

Newdata <- Extra002 %>%
  left_join(Extra003, c("id" = "id_extra")) %>%
   mutate(Price_extra = ifelse(is.na(Price_extra), 0, Price_extra)) %>%
    mutate(New_price = Price_extra + Price01)
View(Newdata)

```



# Problem 7

Make a summary of the purchases for each employee for each day.

Remember to make comments.  

```{r problem 7 solution, eval=FALSE, include=FALSE}

Newdata01 <- Newdata %>% 
  filter(!trimws(Initials) %in% c(NA, "-", "", ".")) %>% 
   group_by(Date01, Initials) %>% 
    dplyr::summarise(
              Count = n(),
               Price = sum(Price01), 
                Price_sub =sum(Price_extra), 
                 Real_price = sum(New_price)
             )



```

# Problem 8

Based on Extra001, keep observations having “product” = “Pant retur” (i.e. bottle deposits). Make a summary per week – i.e. how many observations per week, and how much money is paid back to the employees in total per week.

Remember to make comments.  

```{r problem 8 solution, eval=FALSE, include=FALSE}

Pant_retur <- Extra001 %>% 
  filter(trimws(Product) == "Pant retur") %>% 
   group_by(week(Date01), Initials) %>%
    dplyr::summarize(
              count = n(),
               Pant = sum(Price01)
              )
View(Pant_retur)


```

# Problem 9

Are employees that return the bottles and get deposits back higher spenders in the canteen compared to the rest of the employees?

Remember to make comments.  

```{r problem 9 solution, eval=FALSE, include=FALSE}

Pant_retur01 <- Extra001 %>% 
  filter(trimws(Product) == "Pant retur") %>% 
   distinct(Initials) %>%
    mutate(Pant01 = 1)
View(Pant_retur01)

Newdata02 <- Newdata01 %>% 
  group_by(Initials) %>%
   dplyr::summarize(Total_Sum = sum(Real_price)) 
View(Newdata02)

Newdata03 <- Newdata02 %>% 
  left_join(Pant_retur01, by = "Initials") %>% 
   mutate(Pant01 = ifelse(is.na(Pant01), 0, Pant01)) %>% 
    group_by(Pant01) %>% 
     dplyr::summarize(Average = mean(Total_Sum))
View(Newdata03)

```

# Problem 10

On which weekday (on average) is the sale highest in the canteen?

Remember to make comments.  

```{r problem 10 solution, eval=FALSE, include=FALSE}

Newdata04a <- Newdata01 %>% 
  group_by(weekday = wday(Date01)) %>% 
   dplyr::summarize(Average = mean(Real_price)) %>% 
    arrange(desc(Average)) %>%
     head(1) %>% 
      dplyr::select(weekday)
Newdata04a

Newdata04b <- Newdata01 %>% 
  group_by(weekday = weekdays(Date01)) %>% 
   dplyr::summarize(Total = sum(Real_price)) %>% 
    arrange(desc(Total)) %>%
     head(1) %>% 
      dplyr::select(weekday)
Newdata04b

```

# Problem 11

Calculate the average sales per day but only for employees that are also in sample04, and not in the sample03.

Remember to make comments.  

```{r problem 11 solution, eval=FALSE, include=FALSE}

sample03 <- readRDS("data/sample03.Rda")
sample04 <- readRDS("data/sample04.Rda")

Newdata05 <- Newdata01 %>% 
 mutate(Initials = trimws(Initials)) %>% 
  ungroup() %>% 
   dplyr::select(-(Count:Price_sub))

str(Newdata05)

Newdata05 %>% 
  filter(Initials %in% sample04$Initials & !Initials %in% sample03$Initials) %>% 
   group_by(Date01) %>% 
    dplyr::summarize(
              Average = mean(Real_price),
               Total = sum(Real_price),
                Count = n()
              )

```


# Problem 12

Append the six data sets x201201 to x201206, and call the new data set x2012halfyear.

Remember to make comments.  

```{r problem 12 solution, eval=FALSE, include=FALSE}

x201201 <- readRDS("data/x201201.Rda") 
x201202 <- readRDS("data/x201202.Rda") 
x201203 <- readRDS("data/x201203.Rda") 
x201204 <- readRDS("data/x201204.Rda") 
x201205 <- readRDS("data/x201205.Rda") 
x201206 <- readRDS("data/x201206.Rda") 


# Rigtige, men træls:
x2012halfyear01 <- dplyr::union_all(x201201, dplyr::union_all(x201202, dplyr::union_all(x201203, dplyr::union_all(x201204, dplyr::union_all(x201205, x201206))))) %>% 
  arrange(Dato, Timofday03, Initialer)
x2012halfyear01


# Forkerte:
x2012halfyear01 <- dplyr::union_all(x201201, x201202, x201203, x201204, x201205, x201206) %>% 
  arrange(Dato, Timofday03, Initialer)
x2012halfyear01


# Rigtige:
x2012halfyear02 <- base::rbind(x201201, x201202, x201203, x201204, x201205, x201206) %>% 
  arrange(Dato, Timofday03, Initialer)
x2012halfyear02 

```



# Problem 13

Merge Extra002 and Extra003. Use SQL!

Remember to make comments.  

```{r problem 13 solution, eval=FALSE, include=FALSE}

if (!dir.exists("ourdata")) dir.create("ourdata")

connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")

dbWriteTable(connection_ourdata, "Extra001", Extra001, row.names = FALSE, overwrite = TRUE)
dbWriteTable(connection_ourdata, "Extra002", Extra002, row.names = FALSE, overwrite = TRUE)
dbWriteTable(connection_ourdata, "Extra003", Extra003, row.names = FALSE, overwrite = TRUE)
dbDisconnect(connection_ourdata)


connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")
dbListFields(connection_ourdata, "Extra001")
dbListFields(connection_ourdata, "Extra002")
dbListFields(connection_ourdata, "Extra003")
dbDisconnect(connection_ourdata)

connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")
EXTRA2and3_db  <- dbGetQuery(connection_ourdata, 
                             "select s.*, t.*
                                from Extra002 as s 
                                left join Extra003 as t 
                                on s.id = t.id_extra
                             ")
dbDisconnect(connection_ourdata)

View(EXTRA2and3_db)   


#Extra
connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")
db_Extra001 <- dbReadTable(connection_ourdata, "Extra001")
db_Extra002 <- dbReadTable(connection_ourdata, "Extra002")
db_Extra003 <- dbReadTable(connection_ourdata, "Extra003")
dbDisconnect(connection_ourdata)

```

# Problem 14

Based on Extra001, keep observations having “product” = “Pant retur” (i.e. bottle deposits). Make a summary for each employee – i.e. how many observations per employee, and how much money is paid back to 
the employees in total. Keep only observations if the sum of Price01 is less than -10. Use SQL!  

Remember to make comments.  

```{r problem 14 solution, eval=FALSE, include=FALSE}

connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")
EXTRA_deposit_db <- dbGetQuery(connection_ourdata, 
                               "select *, count(*) as no, sum(Price01) as pr
                                from Extra001
                                 where trim(product) = 'Pant retur'
                                  group by Initials
                                   having pr < -10"
                               )
dbDisconnect(connection_ourdata)
```

```{r}

```

